---
title: "TÜİK Forecasting Project: Number of Visitors by Month (Tourism)"
author: "Emir ARMAN - 138721522"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 5)
```

# Project Overview

This project forecasts the **monthly number of visitors departing from Turkey**, using data obtained directly from the TÜİK Data Portal via the `tuikr` R package and the `httr` package. The forecast target is **April 2026**, the next period after the latest available observation (March 2026).

---

# Data Source and TÜİK Connection

| Field | Information |
|---|---|
| TÜİK Data Set Name | Tourism expenditures, number of visitors and average expenditure per capita by months |
| TÜİK Theme | Tourism (Theme ID: 14) |
| Selected Variable | Number of Visitors (Vatandaş Sayısı) |
| Data Frequency | Monthly |
| Time Coverage | January 2012 – March 2026 |
| Latest Available Observation | March 2026 |
| Forecast Target Period | April 2026 |
| Data Access Method | `tuikr` + `httr::GET()` |
| Package Source | https://github.com/emraher/tuikr |

---

# Data Access via tuikr

```{r data-access}
# Load required packages
library(tuikr)
library(httr)
library(readxl)
library(dplyr)
library(tidyr)
library(forecast)
library(ggplot2)
library(knitr)
library(kableExtra)

# Step 1: Access TÜİK Tourism tables using tuikr
tables14 <- statistical_tables("14")

# Step 2: Get the URL for "Tourism expenditures, number of visitors and average expenditure per capita by months"
url_istab <- tables14$table_url[13]

# Step 3: Download the data using httr::GET() with appropriate headers
# This is necessary because TÜİK's SDMX API (nsiws.tuik.gov.tr) returns HTTP 401.
# The tuikr statistical_tables() function provides istab-type URLs accessible via httr.
temp_file <- tempfile(fileext = ".xls")

response <- GET(url_istab,
                add_headers(
                  "User-Agent" = "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36",
                  "Accept"     = "application/vnd.openxmlformats-officedocument.spreadsheetml.sheet",
                  "Referer"    = "https://veriportali.tuik.gov.tr/"
                ),
                write_disk(temp_file, overwrite = TRUE))

cat("HTTP Status:", status_code(response), "\n")

# Step 4: Read the downloaded xls file
df_raw <- read_xls(temp_file)

cat("Raw data dimensions:", dim(df_raw), "\n")
```

---

# Use of TÜİK Data in R

```{r data-cleaning}
# Extract years from row 3 (every 3rd column starting from col 2)
years <- as.character(df_raw[3, seq(2, 46, by = 3)])
years <- years[!is.na(years)]
years_clean <- gsub("\\(1\\)", "", years)  # Remove "(1)" footnote from 2020

# Visitor count columns: 2nd column of each year group (cols 3, 6, 9, ...)
visitor_cols <- seq(3, 46, by = 3)

# Build tidy data frame: rows 6-17 are January through December
ts_data <- data.frame()

for (i in 1:length(years_clean)) {
  col_idx <- visitor_cols[i]
  values  <- as.numeric(as.character(unlist(df_raw[6:17, col_idx])))
  temp    <- data.frame(
    year     = as.integer(years_clean[i]),
    month    = 1:12,
    visitors = values
  )
  ts_data <- rbind(ts_data, temp)
}

# Add date column
ts_data$date <- as.Date(paste(ts_data$year, sprintf("%02d", ts_data$month), "01", sep = "-"))

# Remove NA rows (future months not yet published)
ts_clean <- ts_data[!is.na(ts_data$visitors), ]

cat("Total observations:", nrow(ts_clean), "\n")
cat("Period:", format(min(ts_clean$date), "%B %Y"), "to", format(max(ts_clean$date), "%B %Y"), "\n")
cat("Latest available observation:", format(max(ts_clean$date), "%B %Y"), "\n")
cat("Forecast target period: April 2026\n")
cat("Data access date:", format(Sys.Date()), "\n")
```

---

# Exploratory Time Series Analysis

```{r eda}
# Create ts object
ts_visitors <- ts(ts_clean$visitors, start = c(2012, 1), frequency = 12)

# Summary statistics
summary(ts_visitors)
```

```{r plot-actual, fig.cap="Actual Time Series: Monthly Number of Visitors"}
# Plot actual series
plot(ts_visitors,
     main = "Monthly Number of Visitors Departing from Turkey (2012–2026)",
     ylab = "Number of Visitors",
     xlab = "Year",
     col  = "steelblue",
     lwd  = 1.5)
grid()
```

```{r decomposition-check}
# STL decomposition for EDA
stl_fit <- stl(ts_visitors, s.window = "periodic", robust = TRUE)
plot(stl_fit, main = "STL Decomposition (EDA)")
```

**Observations:**

- **Trend**: A clear upward trend from 2012 to 2019, a sharp drop in 2020 (COVID-19), and a strong recovery from 2021 onward.
- **Seasonality**: Strong and consistent seasonal pattern — visitor numbers peak in summer months (July–August) and dip in winter.
- **Structural break**: The pandemic in 2020 caused an extraordinary disruption, which is a known structural break.
- **Missing values**: None in the selected period.
- **Outliers**: 2020 values are extreme outliers due to COVID-19 travel restrictions.

---

# Forecasting Methods

We use the **last 36 observations** (2023 Jan – 2026 Mar) as the evaluation window to compute accuracy measures, since the post-COVID recovery period is more representative of current dynamics.

```{r eval-setup}
# Full series
n_total <- length(ts_visitors)

# Use last 36 observations for accuracy evaluation
n_eval  <- 36
ts_eval <- tail(ts_visitors, n_eval)

# Actual values for accuracy calculation
actual <- as.numeric(ts_eval)
n      <- length(actual)
```

---

## Naïve Forecasting

```{r naive}
# Naïve: forecast = previous period value
naive_fit    <- naive(ts_visitors, h = 1)
naive_next   <- as.numeric(naive_fit$mean[1])
naive_fitted <- c(NA, actual[-length(actual)])
naive_errors <- actual - naive_fitted
actual_naive <- actual[!is.na(naive_fitted)]
naive_fitted <- naive_fitted[!is.na(naive_fitted)]
naive_errors <- naive_errors[!is.na(naive_errors)]

# Plot
plot(ts_eval, main = "Naïve Forecast vs Actual (Last 36 Obs)",
     ylab = "Visitors", xlab = "Time", col = "steelblue", lwd = 1.5)
lines(ts(naive_fitted, start = c(2023, 1), frequency = 12),
      col = "red", lwd = 1.5, lty = 2)
legend("topleft", legend = c("Actual", "Naïve Forecast"),
       col = c("steelblue", "red"), lty = c(1, 2), lwd = 2)
grid()
```

---

## Moving Average

```{r moving-average}
# Window size = 12 (one full season) — captures seasonal cycle
ma_window  <- 12
ma_fitted  <- stats::filter(ts_eval, rep(1/ma_window, ma_window), sides = 1)
ma_fitted  <- as.numeric(ma_fitted)

# Remove NAs from the beginning
valid_idx    <- !is.na(ma_fitted)
ma_actual    <- actual[valid_idx]
ma_forecasts <- ma_fitted[valid_idx]
ma_errors    <- ma_actual - ma_forecasts

# Next period forecast: MA of last 12 observations
ma_next <- mean(tail(as.numeric(ts_visitors), 12))

# Plot
plot(ts_eval, main = "Moving Average (k=12) vs Actual",
     ylab = "Visitors", xlab = "Time", col = "steelblue", lwd = 1.5)
lines(ts(ma_fitted, start = c(2023, 1), frequency = 12),
      col = "darkgreen", lwd = 1.5, lty = 2)
legend("topleft", legend = c("Actual", "MA(12) Forecast"),
       col = c("steelblue", "darkgreen"), lty = c(1, 2), lwd = 2)
grid()
```

**Window size justification**: A window of 12 is chosen to match the monthly seasonal cycle, ensuring the moving average captures one full year of variation.

---

## Weighted Moving Average

```{r wma}
# Weights: more recent observations get higher weight
# Last 12 months, linearly increasing weights
wma_window  <- 12
raw_weights <- 1:wma_window
wma_weights <- raw_weights / sum(raw_weights)

wma_fitted <- rep(NA, n)
for (i in wma_window:n) {
  wma_fitted[i] <- sum(wma_weights * actual[(i - wma_window + 1):i])
}

valid_idx2   <- !is.na(wma_fitted)
wma_actual   <- actual[valid_idx2]
wma_forecast <- wma_fitted[valid_idx2]
wma_errors   <- wma_actual - wma_forecast

# Next period forecast
last12       <- tail(as.numeric(ts_visitors), 12)
wma_next     <- sum(wma_weights * last12)

# Plot
plot(ts_eval, main = "Weighted Moving Average (k=12) vs Actual",
     ylab = "Visitors", xlab = "Time", col = "steelblue", lwd = 1.5)
lines(ts(wma_fitted, start = c(2023, 1), frequency = 12),
      col = "purple", lwd = 1.5, lty = 2)
legend("topleft", legend = c("Actual", "WMA(12) Forecast"),
       col = c("steelblue", "purple"), lty = c(1, 2), lwd = 2)
grid()
```

**Weighting logic**: Linearly increasing weights give more importance to recent observations, which is appropriate given the recovery trend post-2021.

---

## Exponential Smoothing

```{r exp-smoothing}
# Simple exponential smoothing — alpha optimized automatically
es_fit    <- ses(ts_visitors, h = 1)
es_alpha  <- es_fit$model$par["alpha"]
cat("Optimal alpha:", round(es_alpha, 4), "\n")

es_fitted <- as.numeric(fitted(es_fit))
es_eval   <- tail(es_fitted, n_eval)
es_errors <- actual - es_eval
es_next   <- as.numeric(es_fit$mean[1])

# Plot
plot(ts_eval, main = paste0("Exponential Smoothing (alpha=", round(es_alpha,3), ") vs Actual"),
     ylab = "Visitors", xlab = "Time", col = "steelblue", lwd = 1.5)
lines(ts(es_eval, start = c(2023, 1), frequency = 12),
      col = "orange", lwd = 1.5, lty = 2)
legend("topleft", legend = c("Actual", "ES Forecast"),
       col = c("steelblue", "orange"), lty = c(1, 2), lwd = 2)
grid()
```

---

## Trend-Adjusted Exponential Smoothing (Holt's Method)

```{r holt}
# Holt's method: captures level + trend
holt_fit    <- holt(ts_visitors, h = 1)
holt_alpha  <- holt_fit$model$par["alpha"]
holt_beta   <- holt_fit$model$par["beta"]
cat("Alpha:", round(holt_alpha, 4), "| Beta:", round(holt_beta, 4), "\n")

holt_fitted <- as.numeric(fitted(holt_fit))
holt_eval   <- tail(holt_fitted, n_eval)
holt_errors <- actual - holt_eval
holt_next   <- as.numeric(holt_fit$mean[1])

# Plot
plot(ts_eval, main = paste0("Holt's Trend-Adjusted ES (alpha=", round(holt_alpha,3),
                             ", beta=", round(holt_beta,3), ") vs Actual"),
     ylab = "Visitors", xlab = "Time", col = "steelblue", lwd = 1.5)
lines(ts(holt_eval, start = c(2023, 1), frequency = 12),
      col = "brown", lwd = 1.5, lty = 2)
legend("topleft", legend = c("Actual", "Holt Forecast"),
       col = c("steelblue", "brown"), lty = c(1, 2), lwd = 2)
grid()
```

**Trend assessment**: The series shows a clear upward trend post-2021. Holt's method is therefore applicable.

---

## Linear Trend Projection

```{r linear-trend}
# Time index
t_index   <- 1:n_total
lm_fit    <- lm(as.numeric(ts_visitors) ~ t_index)

cat("Intercept:", round(coef(lm_fit)[1], 2), "\n")
cat("Slope:", round(coef(lm_fit)[2], 2), "\n")

lm_fitted  <- as.numeric(fitted(lm_fit))
lm_eval    <- tail(lm_fitted, n_eval)
lm_errors  <- actual - lm_eval

# Forecast next period (t = n_total + 1)
lm_next    <- predict(lm_fit, newdata = data.frame(t_index = n_total + 1))

# Plot
plot(ts_eval, main = "Linear Trend Projection vs Actual",
     ylab = "Visitors", xlab = "Time", col = "steelblue", lwd = 1.5)
lines(ts(lm_eval, start = c(2023, 1), frequency = 12),
      col = "darkred", lwd = 1.5, lty = 2)
legend("topleft", legend = c("Actual", "Linear Trend"),
       col = c("steelblue", "darkred"), lty = c(1, 2), lwd = 2)
grid()
```

**Interpretation**: The intercept represents the estimated visitor count at time zero (extrapolated baseline). The slope indicates the average monthly increase in visitor numbers over the study period.

---

## Seasonal Indices

```{r seasonal-indices}
# Calculate seasonal indices using ratio-to-moving-average method
# Step 1: 12-month centered moving average
ma12       <- stats::filter(as.numeric(ts_visitors), rep(1/12, 12), sides = 2)

# Step 2: Ratio of actual to MA
ratios     <- as.numeric(ts_visitors) / ma12

# Step 3: Average ratio by month = seasonal index
months_vec <- cycle(ts_visitors)
si         <- tapply(ratios, months_vec, mean, na.rm = TRUE)

# Normalize so indices sum to 12
si         <- si * (12 / sum(si))

cat("Seasonal Indices:\n")
month_names <- c("Jan","Feb","Mar","Apr","May","Jun",
                 "Jul","Aug","Sep","Oct","Nov","Dec")
si_df <- data.frame(Month = month_names, Seasonal_Index = round(si, 4))
print(si_df)

# Deseasonalized series
deseason   <- as.numeric(ts_visitors) / si[months_vec]

# Trend on deseasonalized
t_index2   <- 1:n_total
lm_deseas  <- lm(deseason ~ t_index2)
trend_vals <- predict(lm_deseas, newdata = data.frame(t_index2 = 1:n_total))

# Reseasonalize
si_fitted  <- trend_vals * si[months_vec]
si_eval    <- tail(si_fitted, n_eval)
si_errors  <- actual - si_eval

# Next period forecast (April = month 4)
next_t     <- n_total + 1
next_trend <- predict(lm_deseas, newdata = data.frame(t_index2 = next_t))
si_next    <- next_trend * si[4]  # April index

# Plot
plot(ts_eval, main = "Seasonal Indices Forecast vs Actual",
     ylab = "Visitors", xlab = "Time", col = "steelblue", lwd = 1.5)
lines(ts(si_eval, start = c(2023, 1), frequency = 12),
      col = "darkgreen", lwd = 1.5, lty = 2)
legend("topleft", legend = c("Actual", "Seasonal Indices"),
       col = c("steelblue", "darkgreen"), lty = c(1, 2), lwd = 2)
grid()
```

**High season**: Months with index > 1 (July, August) indicate above-average visitor counts. **Low season**: Months with index < 1 (January, February) indicate below-average counts.

---

## Additive Decomposition

```{r additive-decomp}
# Classical additive decomposition
add_decomp  <- decompose(ts_visitors, type = "additive")
plot(add_decomp)

# Reconstruct fitted values (trend + seasonal)
add_fitted  <- as.numeric(add_decomp$trend) + as.numeric(add_decomp$seasonal)
add_eval    <- tail(add_fitted[!is.na(add_fitted)], n_eval)
add_actual2 <- tail(actual, length(add_eval))
add_errors  <- add_actual2 - add_eval

# Forecast next period
last_trend  <- tail(as.numeric(add_decomp$trend)[!is.na(add_decomp$trend)], 1)
add_next    <- last_trend + as.numeric(add_decomp$seasonal)[4]  # April seasonal component

# Plot
plot(ts(add_actual2, start = c(2023, 1), frequency = 12),
     main = "Additive Decomposition Forecast vs Actual",
     ylab = "Visitors", xlab = "Time", col = "steelblue", lwd = 1.5)
lines(ts(add_eval, start = c(2023, 1), frequency = 12),
      col = "magenta", lwd = 1.5, lty = 2)
legend("topleft", legend = c("Actual", "Additive Decomp"),
       col = c("steelblue", "magenta"), lty = c(1, 2), lwd = 2)
grid()
```

---

## Multiplicative Decomposition

```{r multiplicative-decomp}
# Classical multiplicative decomposition
mul_decomp  <- decompose(ts_visitors, type = "multiplicative")
plot(mul_decomp)

# Reconstruct fitted values (trend * seasonal)
mul_fitted  <- as.numeric(mul_decomp$trend) * as.numeric(mul_decomp$seasonal)
mul_eval    <- tail(mul_fitted[!is.na(mul_fitted)], n_eval)
mul_actual2 <- tail(actual, length(mul_eval))
mul_errors  <- mul_actual2 - mul_eval

# Forecast next period
last_trend2 <- tail(as.numeric(mul_decomp$trend)[!is.na(mul_decomp$trend)], 1)
mul_next    <- last_trend2 * as.numeric(mul_decomp$seasonal)[4]

# Plot
plot(ts(mul_actual2, start = c(2023, 1), frequency = 12),
     main = "Multiplicative Decomposition Forecast vs Actual",
     ylab = "Visitors", xlab = "Time", col = "steelblue", lwd = 1.5)
lines(ts(mul_eval, start = c(2023, 1), frequency = 12),
      col = "darkcyan", lwd = 1.5, lty = 2)
legend("topleft", legend = c("Actual", "Multiplicative Decomp"),
       col = c("steelblue", "darkcyan"), lty = c(1, 2), lwd = 2)
grid()
```

---

## Regression with Trend and Seasonal Dummy Variables

```{r regression-seasonal}
# Create data frame with time index and seasonal dummies
reg_df           <- data.frame(
  visitors = as.numeric(ts_visitors),
  t        = 1:n_total,
  month    = factor(cycle(ts_visitors))
)

# Fit regression (month 1 = January is the reference category)
reg_fit    <- lm(visitors ~ t + month, data = reg_df)
summary(reg_fit)

reg_fitted <- as.numeric(fitted(reg_fit))
reg_eval   <- tail(reg_fitted, n_eval)
reg_errors <- actual - reg_eval

# Forecast April 2026 (t = n_total + 1, month = 4)
reg_next   <- predict(reg_fit,
                      newdata = data.frame(t = n_total + 1,
                                           month = factor(4, levels = 1:12)))

# Plot
plot(ts_eval, main = "Regression (Trend + Seasonal Dummies) vs Actual",
     ylab = "Visitors", xlab = "Time", col = "steelblue", lwd = 1.5)
lines(ts(reg_eval, start = c(2023, 1), frequency = 12),
      col = "navy", lwd = 1.5, lty = 2)
legend("topleft", legend = c("Actual", "Regression Forecast"),
       col = c("steelblue", "navy"), lty = c(1, 2), lwd = 2)
grid()
```

**Trend coefficient**: Represents the average monthly increase in visitor numbers after controlling for seasonality. **Seasonal dummies**: Each month coefficient shows the deviation from January (reference month).

---

# Forecast Accuracy Comparison

```{r accuracy-function}
# Function to compute all required accuracy measures
calc_accuracy <- function(actual, forecast, method_name) {
  e    <- actual - forecast
  n    <- length(e)
  bias <- mean(e)
  mad  <- mean(abs(e))
  mse  <- mean(e^2)
  mape <- mean(abs(e / actual)) * 100
  rsfe <- sum(e)
  ts_  <- rsfe / mad  # Tracking Signal
  data.frame(
    Method          = method_name,
    Bias            = round(bias, 2),
    MAD             = round(mad, 2),
    MSE             = round(mse, 2),
    MAPE            = round(mape, 4),
    RSFE            = round(rsfe, 2),
    Tracking_Signal = round(ts_, 4),
    stringsAsFactors = FALSE
  )
}
```

```{r accuracy-table}
# Compute accuracy for each method
# Note: additive and multiplicative decomp may have fewer eval points due to NA trend edges
acc_naive <- calc_accuracy(actual,       naive_errors + actual - naive_errors,  "Naïve")
# Recalculate properly
acc_naive <- calc_accuracy(actual,                  actual - naive_errors,  "Naïve")
acc_ma    <- calc_accuracy(ma_actual,    ma_forecasts,   "Moving Average (k=12)")
acc_naive <- calc_accuracy(actual_naive, naive_fitted, "Naïve")
acc_wma   <- calc_accuracy(wma_actual,   wma_forecast,   "Weighted Moving Average (k=12)")
acc_es    <- calc_accuracy(actual,       actual - es_errors,    "Exponential Smoothing")
acc_lm    <- calc_accuracy(actual,       actual - lm_errors,    "Linear Trend Projection")
acc_si    <- calc_accuracy(actual,       actual - si_errors,    "Seasonal Indices")
acc_add   <- calc_accuracy(add_actual2,  add_eval,              "Additive Decomposition")
acc_mul   <- calc_accuracy(mul_actual2,  mul_eval,              "Multiplicative Decomposition")
acc_reg   <- calc_accuracy(actual,       actual - reg_errors,   "Regression Trend+Seasonal")

# Next period forecasts
next_forecasts <- c(
  round(naive_next, 0),
  round(ma_next, 0),
  round(wma_next, 0),
  round(es_next, 0),
  round(holt_next, 0),
  round(lm_next, 0),
  round(si_next, 0),
  round(add_next, 0),
  round(mul_next, 0),
  round(as.numeric(reg_next), 0)
)

# Combine all accuracy results
acc_all <- rbind(acc_naive, acc_ma, acc_wma, acc_es, acc_holt,
                 acc_lm, acc_si, acc_add, acc_mul, acc_reg)
acc_all$Next_Period_Forecast_Apr2026 <- next_forecasts

# Display table
kable(acc_all, caption = "Forecast Accuracy Comparison — All Methods",
      format = "html", align = "c") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 12)
```

```{r save-accuracy-table}
# Save accuracy table
dir.create("outputs/tables", recursive = TRUE, showWarnings = FALSE)
write.csv(acc_all, "outputs/tables/accuracy_comparison.csv", row.names = FALSE)
cat("Accuracy table saved.\n")
```

---

# Superior Method Selection

```{r superior-method}
# Find method with lowest MAPE
best_idx    <- which.min(acc_all$MAPE)
best_method <- acc_all$Method[best_idx]
cat("Method with lowest MAPE:", best_method, "\n")
cat("MAPE:", acc_all$MAPE[best_idx], "%\n")
```

**Selection Justification:**

The series exhibits both a clear **upward trend** (post-2021 recovery) and strong **monthly seasonality**. Methods that capture both components simultaneously are most suitable.

- **Regression with Trend and Seasonal Dummies** directly models both trend and seasonality through the time index and dummy variables — it is structurally the most appropriate method for this data.
- The **Multiplicative Decomposition** is also strong, since the seasonal amplitude grows with the level of the series (multiplicative pattern).
- The tracking signal for the selected method should be within the ±4 MAD control limits, indicating no systematic bias.
- Simple methods (Naïve, MA, WMA) do not capture the seasonal structure adequately.
- Exponential Smoothing without seasonality is also limited for this highly seasonal series.

**Selected superior method: Regression with Trend and Seasonal Dummy Variables**

---

# Final Next-Period Forecast

```{r final-forecast}
# Final forecast for April 2026 using Regression (Trend + Seasonal Dummies)
final_value  <- round(as.numeric(reg_next), 0)
final_method <- "Regression with Trend and Seasonal Dummy Variables"
access_date  <- format(Sys.Date(), "%Y-%m-%d")

cat("=== FINAL FORECAST ===\n")
cat("Selected Method       :", final_method, "\n")
cat("Data Access Date      :", access_date, "\n")
cat("Latest TÜİK Observation: March 2026\n")
cat("Forecast Target Period : April 2026\n")
cat("Forecasted Value      :", format(final_value, big.mark = ","), "visitors\n")
```

```{r final-plot, fig.cap="Superior Method: Actual vs Fitted + April 2026 Forecast"}
# Plot showing fitted values and next-period forecast
ts_fitted_full <- ts(reg_fitted, start = c(2012, 1), frequency = 12)

# April 2026 point
apr2026_date <- as.Date("2026-04-01")

plot(ts_visitors,
     main = "Regression (Trend + Seasonal Dummies): Actual vs Fitted + April 2026 Forecast",
     ylab = "Number of Visitors",
     xlab = "Year",
     col  = "steelblue",
     lwd  = 1.5)
lines(ts_fitted_full, col = "red", lwd = 1.5, lty = 2)
points(2026 + 3/12, final_value, pch = 19, col = "darkred", cex = 2)
text(2026 + 3/12, final_value,
     labels = paste0("Apr 2026\n", format(final_value, big.mark = ",")),
     pos = 3, col = "darkred", cex = 0.85)
legend("topleft",
       legend = c("Actual", "Regression Fitted", "April 2026 Forecast"),
       col    = c("steelblue", "red", "darkred"),
       lty    = c(1, 2, NA), pch = c(NA, NA, 19), lwd = 2)
grid()
```

```{r save-final-forecast}
# Save final forecast
final_df <- data.frame(
  Superior_Method          = final_method,
  Data_Access_Date         = access_date,
  Latest_Available_Obs     = "March 2026",
  Forecast_Target_Period   = "April 2026",
  Forecasted_Value         = final_value,
  MAPE                     = acc_all$MAPE[acc_all$Method == "Regression Trend+Seasonal"],
  MAD                      = acc_all$MAD[acc_all$Method  == "Regression Trend+Seasonal"],
  Tracking_Signal          = acc_all$Tracking_Signal[acc_all$Method == "Regression Trend+Seasonal"]
)

write.csv(final_df, "outputs/tables/final_forecast.csv", row.names = FALSE)
cat("Final forecast saved.\n")
```

---

# Interpretation of Results

The regression model with trend and seasonal dummy variables forecasts approximately **`r format(final_value, big.mark=",")`** visitors departing from Turkey in **April 2026**.

April is typically a transitional month for Turkish outbound tourism — higher than winter months but below the summer peak. The seasonal index for April is above 1, reflecting the beginning of the spring travel surge. The forecast is consistent with the recovery trajectory observed since 2021 and aligns with the seasonal pattern established over the 2012–2019 baseline.

---

# Limitations

1. **Structural break (COVID-19)**: The 2020 data represents an extraordinary disruption. Including these observations may distort trend and seasonal estimates.
2. **Short post-recovery period**: Only 3–4 years of post-COVID data are available, which may not fully capture the new baseline dynamics.
3. **No external variables**: The model does not include macroeconomic indicators, exchange rates, or geopolitical factors that influence outbound tourism.
4. **Linear trend assumption**: The regression assumes a constant linear trend, which may not hold if growth accelerates or decelerates.
5. **Data revisions**: TÜİK may revise earlier figures, which could affect accuracy measures retrospectively.

---

# Reproducibility

To reproduce this analysis:

1. Install R (>= 4.6) and RStudio
2. Clone the repository
3. Install `renv` and run `renv::restore()` to restore the package environment
4. Open `forecasting_project.Rmd` and click **Knit**

All data are accessed programmatically via `tuikr` and `httr::GET()`. No manual data files are used.

---

# Repository Structure

```
tuik-forecasting-project/
├── README.md
├── forecasting_project.Rmd
├── forecasting_project.html
├── outputs/
│   ├── tables/
│   │   ├── accuracy_comparison.csv
│   │   └── final_forecast.csv
│   └── figures/
├── R/
│   ├── data_import.R
│   ├── forecasting_methods.R
│   ├── accuracy_measures.R
│   └── plots.R
├── renv.lock
└── .gitignore
```

---

*Project completed by Emir ARMAN (138721522) — Forecasting Course*
